---
title: "Model recovery"
author: "LV"
date: "2024-01-10"
output: html_document
---

Script takes model recovery simulation outputs and checks for model recovery, 
Plots confusion matrix for

A) Recovery: Value vs. perceptual 
B) Reduced parametrization: Gauss Value (alpha =0) vs. perceptual 

# Get environment right
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("renv", "here", "knitr", "dplyr")

here::i_am("flag_project_root.R")
here::here()
data_path = here::here()

#renv::activate(project=here::here())
#renv::restore(project=here::here()) #when asked to activate: NO!
#renv::snapshot()

packages <- c("ggplot2", "dplyr", "stats","lme4", "here", "renv","gtools", "see","plotrix","modelbased","emmeans", "performance", "tidyverse", "ggsignif","broom","quickpsy", "ggpubr", "performance","glmmTMB", "loo", "bayesplot","rlang","bayestestR", "rstan","brms","tidybayes","sjPlot","sjmisc","yardstick","effectsize","datawizard","esc","grid","lsr","gridExtra")

# load the packages from lock 
lapply(packages, library, character.only = TRUE)
```


#Model recovery perceptual vs value model 

# Get and prep data
```{r}
## Define paths 
outs = here::here('outputs', 'model_recovery','ITER30LINONE') #change in bounds
outs2 = here::here('outputs', 'model_recovery','ITER30GAUSSONE') #change in bounds

all_files <- list.files(path = outs, pattern = "\\.csv$", full.names = TRUE)
all_files2 <- list.files(path = outs2, pattern = "\\.csv$", full.names = TRUE)

# Get all files
csv_files <- all_files[grep(".csv", all_files, ignore.case = TRUE) &
                       grepl("recovery", all_files, ignore.case = TRUE)]
csv_files2 <- all_files2[grep(".csv", all_files2, ignore.case = TRUE) &
                       grepl("recovery", all_files, ignore.case = TRUE)]
  
# Read in CSV files into a list of data frames
data_list <- lapply(csv_files, read.csv)
data_list2 <- lapply(csv_files2, read.csv)
df2 <- do.call(rbind, data_list)
df22 <- do.call(rbind, data_list2)
df22$sub = paste0(df22$sub, "X")

df <- rbind(df2,df22)
df2 <- df
df2 <- df2 %>% select(where(~ !all(is.na(.))))
```

# Get cumulative BIC across iterations and and conditions
```{r}
all_cols = c()

# All model combinations of interest
all_combi = c("BIC_perc_est_perc_est", "BIC_perc_est_value_alt_offset", "BIC_value_alt_offset_perc_est", "BIC_value_alt_offset_value_alt_offset")

# Group by subject and combination, then find the minimum for each combination
df2 <- df2 %>%
  group_by(sub) %>%
  mutate(across(all_of(all_combi), ~min(.), .names = "full_{.col}")) %>%
  ungroup()

# Select relevant columns
dff <- df2[, c("full_BIC_perc_est_perc_est", "full_BIC_perc_est_value_alt_offset", "full_BIC_value_alt_offset_perc_est", "full_BIC_value_alt_offset_value_alt_offset","sub","iter")] 

# One row x ppt.
dff = dff[dff$iter == 1, ]
dff$simulated = NA
dff <- dff[, -which(names(dff) == "iter")]

# Pivot and bring to pretty format
dfr <- dff %>%
   select(c("full_BIC_perc_est_perc_est", "full_BIC_perc_est_value_alt_offset", "full_BIC_value_alt_offset_perc_est", "full_BIC_value_alt_offset_value_alt_offset","sub")) %>%
  pivot_longer(cols = c("full_BIC_perc_est_perc_est", "full_BIC_perc_est_value_alt_offset", "full_BIC_value_alt_offset_perc_est", "full_BIC_value_alt_offset_value_alt_offset"), names_to = "model", values_to = "BIC") 

dfr$model = paste0(dfr$model, "X")
dfr$simulated = NA
dfr$simulated[grepl("BIC_perc_est", dfr$model)] <- "perception"
dfr$simulated[grepl("BIC_value_alt_offset", dfr$model)] <- "value_alt_offset"

dfr$fitted = NA
dfr$fitted[grepl("perc_estX", dfr$model)] <- "perception"
dfr$fitted[grepl("value_alt_offsetX", dfr$model)] <- "value_alt_offset"
```


# Check which model fit best depending on simulation
```{r}
dfr$best_fit = NA
dfr <- dfr %>%
  group_by(sub, simulated) %>%
  mutate(best_fit = fitted[which.min(BIC)]) %>%
  ungroup()

# Extract one row per sub/simulated instance
fdf <- dfr %>%
  group_by(sub, simulated) %>%
  slice_min(order_by = BIC, n = 1) %>%
  ungroup()

# Finalize subsets
fdfa <- fdf %>%
  select(simulated, best_fit, BIC)

fdf <- fdf %>%
  select(simulated, best_fit)
```


# Plot
```{r}
# Rename
fdf$simulated[fdf$simulated == "value_alt_offset"] = "value"
fdf$best_fit[fdf$best_fit == "value_alt_offset"] = "value"

# Define levels
possible_levels <- c("perception","value")
fdf$simulated <- factor(fdf$simulated, levels = possible_levels)
fdf$best_fit <- factor(fdf$best_fit,
                 levels = possible_levels)

# Confusion matrix
cm <- yardstick::conf_mat(fdf, simulated, best_fit)

# Plot as heatmap 
ggplot2::autoplot(cm, type = "heatmap") +
  scale_fill_gradient(low = "#ABA8BE", high = "#F5D9C2")+theme(aspect.ratio = 1)+xlab("true model")+ylab("recovered model")+theme(text = element_text(size = 20))

# save
#ggsave(path = paste0(data_path, "/plots"), width = 5, height = 5, filename = 'modelrec.png', device='png', dpi=700) 
```





#Model recovery for reduced parametrization: Gaussian value model (offset fixed at 0) vs. Perceptual model 

```{r}
# Get data 
outs = here::here('outputs', 'model_recovery', 'ITER30ALPHAZERO120')
all_files <- list.files(path = outs, pattern = "\\.csv$", full.names = TRUE)
csv_files <- all_files[grep(".csv", all_files, ignore.case = TRUE) &
                       grepl("recovery", all_files, ignore.case = TRUE)]
# Read in CSV files into a list of data frames
data_list <- lapply(csv_files, read.csv)
df2 <- do.call(rbind, data_list)
df2 <- df2 %>% select(where(~ !all(is.na(.))))
```


# Get cumulative BIC across iterations and and conditions
```{r}
all_cols = c()
all_combi = c("BIC_perc_est_perc_est", "BIC_perc_est_value_alt_offset", "BIC_value_alt_offset_perc_est", "BIC_value_alt_offset_value_alt_offset")

for (s in unique(df2$sub)){
  for (ic in all_combi){
    
  df2temp = subset(df2, sub == s)
  df2s <- df2temp %>% select(ic)
  full_add = min(df2s)
  df2[[paste0("full_", ic)]][df2$sub == s] <- full_add
  }
}

#get relevant columns and one row x ppt
dff <- df2[, c("full_BIC_perc_est_perc_est", "full_BIC_perc_est_value_alt_offset", "full_BIC_value_alt_offset_perc_est", "full_BIC_value_alt_offset_value_alt_offset","sub","iter")] 

dff = dff[dff$iter == 1, ]
dff$simulated = NA
dff <- dff[, -which(names(dff) == "iter")]

dfr <- dff %>%
   select(c("full_BIC_perc_est_perc_est", "full_BIC_perc_est_value_alt_offset", "full_BIC_value_alt_offset_perc_est", "full_BIC_value_alt_offset_value_alt_offset","sub")) %>%
  pivot_longer(cols = c("full_BIC_perc_est_perc_est", "full_BIC_perc_est_value_alt_offset", "full_BIC_value_alt_offset_perc_est", "full_BIC_value_alt_offset_value_alt_offset"), names_to = "model", values_to = "BIC") 

dfr$model = paste0(dfr$model, "X")

dfr$simulated = NA
dfr$simulated[grepl("BIC_perc_est", dfr$model)] <- "perception"
dfr$simulated[grepl("BIC_value_alt_offset", dfr$model)] <- "value_alt_offset"

dfr$fitted = NA
dfr$fitted[grepl("perc_estX", dfr$model)] <- "perception"
dfr$fitted[grepl("value_alt_offsetX", dfr$model)] <- "value_alt_offset"
```


# Check which model fit best depending on simulation
```{r}
dfr$best_fit = NA
dfr <- dfr %>%
  group_by(sub, simulated) %>%
  mutate(best_fit = fitted[which.min(BIC)]) %>%
  ungroup()

# Extract one row per sub/simulated instance
fdf <- dfr %>%
  group_by(sub, simulated) %>%
  slice_min(order_by = BIC, n = 1) %>%
  ungroup()

# Finalize subsets
fdfa <- fdf %>%
  select(simulated, best_fit, BIC)

fdf <- fdf %>%
  select(simulated, best_fit)
```




```{r}
fdf$simulated[fdf$simulated == "value_alt_offset"] = "value"
fdf$best_fit[fdf$best_fit == "value_alt_offset"] = "value"

possible_levels <- c("perception","value")
fdf$simulated <- factor(fdf$simulated, levels = possible_levels)
fdf$best_fit <- factor(fdf$best_fit,
                 levels = possible_levels)

cm <- yardstick::conf_mat(fdf, simulated, best_fit)

autoplot(cm, type = "heatmap") +
  scale_fill_gradient(low = "#ABA8BE", high = "#F5D9C2")+theme(aspect.ratio = 1)+xlab("true model")+ylab("recovered model")+theme(text = element_text(size = 20))

#ggsave(path = paste0(data_path, "/plots/Supplement"), width = 5, height = 5, filename = 'modelrecalpha0.png', device='png', dpi=700) 
```
